---
title: "IMDb classification"
output: html_notebook
---

In this example, we'll try to classify whether movie reviews on IMDb are positive or negative based on the review text.

We'll start by loading some data that comes with the **keras** package. Here, `max_features` denotes the number of unique word tokens we want to consider.

```{r}
library(keras)
max_features <- 20000
imdb <- dataset_imdb(num_words = max_features)
c(c(x_train, y_train), c(x_test, y_test)) %<-% imdb
```

Let's see what the training data looks like

```{r}
x_train[[1]]
```

To turn this into English, we can use the following helper function

```{r}
word_index <- dataset_imdb_word_index()
reverse_word_index <- names(word_index)
names(reverse_word_index) <- word_index
decode_imdb <- function(indexes) {
  force(indexes)
  words <- sapply(indexes, function(index) {
    word <- if (index >= 3) reverse_word_index[[as.character(index - 3)]]
    if (!is.null(word)) word else "?"
  })
  paste(words, collapse = " ")
}
decode_imdb(x_train[[1]])
```

Since the reviews are of variable length, we need to get them all to the same length before we can feed them to the model.

```{r}
maxlen <- 80
x_train <- pad_sequences(x_train, maxlen = maxlen)
x_test <- pad_sequences(x_test, maxlen = maxlen)
```

Our `x_train` is now a matrix where each row corresponds to a review:

```{r}
str(x_train)
```

We can then define a model and compile it:

```{r}
model <- keras_model_sequential()
model %>%
  layer_embedding(input_dim = max_features, output_dim = 128) %>% 
  layer_gru(units = 32, dropout = 0.2, recurrent_dropout = 0.2) %>% 
  layer_dense(units = 1, activation = 'sigmoid')

model %>% compile(
  loss = 'binary_crossentropy',
  optimizer = 'adam',
  metrics = c('accuracy')
)
```

Train the model and evaluate its performance

```{r}
model %>% fit(
  x_train, y_train,
  batch_size = batch_size,
  epochs = 3,
  validation_data = list(x_test, y_test)
)

scores <- model %>% evaluate(
  x_test, y_test,
  batch_size = batch_size
)

cat('Test loss:', scores[[1]])
cat('Test accuracy', scores[[2]])
```

