---
title: "Predicting flight arrival delays"
output: html_notebook
---

For this exercise, we'll work with the `flights` table from the `nycflights13` package. 

```{r}
library(tidyverse)
flights <- nycflights13::flights
glimpse(flights)
```

Our goal in this exercise is to predict arrival delays using information available as of takeoff for each flight.

Let's do some light feature engineering:

```{r}
data <- flights %>%
  filter(!is.na(arr_delay)) %>%
  mutate(orig_dest = paste0(origin, dest)) %T>%
  (function(x) {
    orig_dest_levels <<- unique(x$orig_dest)
    carrier_levels <<- unique(x$carrier)
  }) %>%
  mutate(
    orig_dest = factor(orig_dest, levels = orig_dest_levels) %>%
      as.integer() %>%
      subtract(1),
    carrier = factor(carrier, levels = carrier_levels) %>%
      as.integer() %>%
      subtract(1),
    key = ifelse(month >= 11, "test", "train")
    )
training_data <- filter(data, key == "train")
testing_data <- filter(data, key == "test")
```

We incorporate each predictor we want to include in the model by creating an input layer for it, then concatenating them.

```{r}
library(keras)
carrier_input <- layer_input(shape = c(1))
dep_delay_input <- layer_input(shape = c(1))
distance_input <- layer_input(shape = c(1))
origin_destination_input <- layer_input(shape = c(1))

carrier_output <- carrier_input %>%
  layer_embedding(16, 8) %>%
  layer_flatten()

origin_destination_output <- origin_destination_input %>%
  layer_embedding(256, 128) %>%
  layer_flatten()

main_layer <- layer_concatenate(
  c(dep_delay_input, distance_input,
    carrier_output, origin_destination_output)
  )

output <- main_layer %>%
  layer_dense(units = 256, activation = "relu") %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 256) %>%
  (function(x) layer_concatenate(c(x, dep_delay_input))) %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 128) %>%
  layer_dense(units = 1, activation = "linear")

model <- keras_model(
  list(dep_delay_input, distance_input, carrier_input,
       origin_destination_input),
  output
)
model %>%
  compile(optimizer = "adam", loss = "mse")
```

We're passing our categorical variables through an embedding layer via `layer_embedding()`. The use case here is slightly different from the NLP example, but the intuition is similar: we're asking the model to learn a lower-dimensional representation of the high cardinality categorical variables in the context of the predictive modeling problem.

To make training easier, we're going to scale each of the numeric variables, including the response. Here we define a function factory and use it create some helper functions.

```{r}
make_scaler <- function(x) {
  scaled <- scale(x)
  center_value <- attr(scaled, "scaled:center")
  scale_value <- attr(scaled, "scaled:scale")
  list(
    scaler = function(x) {
      force(x)
      scale(x, center_value, scale_value)
    },
    descaler = function(x) {
      force(x)
      x * scale_value + center_value
    }
  )
}

c(scale_arr_delay, descale_arr_delay) %<-% make_scaler(training_data$arr_delay)
c(scale_distance, descale_distance) %<-% make_scaler(training_data$distance)
c(scale_dep_delay, descale_dep_delay) %<-% make_scaler(training_data$dep_delay)
```

Let's now traing the model!

```{r}
history <- model %>%
  fit(list(scale_dep_delay(training_data$dep_delay), 
           scale_distance(training_data$distance), 
           training_data$carrier, training_data$orig_dest), 
      scale_arr_delay(training_data$arr_delay),
      batch_size = 256, epochs = 10,
      validation_split = 0.2)
```

Score the testing dataset:

```{r}
predictions <- predict(
  model, 
  list(
    scale_dep_delay(testing_data$dep_delay), 
    scale_distance(testing_data$distance), 
    testing_data$carrier, testing_data$orig_dest)
) %>%
  descale_arr_delay()
```

Look at some qualitative performance results:

```{r}
testing_data %>%
  add_column(prediction = as.vector(predictions)) %>%
  select(arr_delay, prediction) %>%
  mutate(decile = cut(prediction, quantile(prediction, probs = seq(0, 1, 0.1)),
                      labels = FALSE, include.lowest = TRUE)
         ) %>%
  group_by(decile) %>%
  summarize(mean_actual = mean(arr_delay),
            mean_predicted = mean(prediction)
            ) %>%
  gather(key, value, mean_actual, mean_predicted) %>%
  ggplot(aes(x = as.factor(decile), y = value, fill = key)) +
  geom_bar(stat = "identity", position = "dodge")
```

